A Large-Scale Study of the Use of Eval in R

### 2020-08-22 inst/paper/main.tex
intro

\newcommand{\eval}{\texttt{eval}\xspace}

\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}

@@

The \c{eval} function has a bad reputation. This function, under one name or
another and with various exact semantics, is available in many dynamic
languages. It is used to transform text into code and execute that code.  It
earned its reputation because of the challenges it poses for reasoning about
code that uses it.  The problem with eval is that the behavior of a system
departs from the source text of the program and becomes dependent on the
program's inputs. This brings challenge to program analysis and security.

>>>>

\newcommand{\eval}{\texttt{eval}\xspace}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}

@@

Traditional program analysis techniques are based on sound reasoning about
program behavior, a program analyzer computes an over-approximation of the
set of possible operations performed by the system being
analyzed~\cite{cc77}.  This is predicated on knowing the program's code, and
computing that code's behavior for all possible inputs. The presence of
\eval entails that the program's input, any string variable, may be
reflected into its code giving rise to \emph{any} behavior allowed by the
language semantics. In many dynamic languages, \eval can redefine most
user-defined and built-in functions resulting in a complete loss of
precision for any operation following \eval. A group of influential
resaerchers argued to give up on soundness and, instead, to
under-approximate dynamic features~\cite{soundy} In their words ``a
practical analysis, therefore, may pretend that \eval does nothing, unless
it can precisely resolve its string argument at compile time.''  Assuming
that \eval does not have side-effects may be too optimistic. We rather
propose to study what \eval does in practice, in real-world code, so as to
give researchers data about how to better under-approximate uses of that
feature.

While there has been previous work investigating the usage of \eval in web
programming and JavaScript, there are no other studies that shed light on
the usage of \eval in other programming languages and application domains.
It is thus reasonable to wonder if the results from previous work can be
carried over to other contexts. For this paper, we have picked the data
science language R~\cite{r} as our object of study. R is both different from
JavaScript in linguistic terms and in terms of its user base. JavaScript was
designed to run untrusted code in browser, while R was design for
statistical computing on a desktop. JavaScript is used as a general purpose
language by a wide community of programmers; while R is used for scientific
computing by data scientist and domain experts with, often, limited
programming experience.

This paper sets out to highlight the differences in usage patterns between
these two languages and user communities. Our results are broadly applicable
as they increase our understanding of the range of uses of \eval, present
new patterns, and hopefully provide insights into reasonable
under-approximations of \eval. More narrowly, for researchers interested in
static analysis of data science code, we provide a publicly available corpus
of programs and a toolchain that can be customized to deliver different
types of information about the behavior of \eval.
